---
title: "Heart eQTLs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(liftOver)
library(ggpubr)
library(rstatix)
source('../R/analysis_utils.R')
```


### Heart eQTL Finemapping

Here we focus on eQTLs that are finemapped with a posterior probability of association > 0.8. There are approximately 2,300 such SNPs.

```{r}
finemap.res <- suppressMessages(readr::read_tsv('../eQTL_enrich/gtex_finemapping/GTEx_v8_finemapping_DAPG/Heart_LV_Finemapping_CS95.txt', col_names = F))
colnames(finemap.res) <- c("tissue","gene","cluster_id","cluster_pip","variant_id","variant_pip")
```

```{r}
high_pip_eqtls <- finemap.res[finemap.res$variant_pip>0.8,] %>% group_by(variant_id) %>% arrange(-variant_pip) %>% slice(1) %>% select(variant_id, variant_pip, gene)
eqtl.gr <- snpIDtoGR(high_pip_eqtls$variant_id)
eqtl.gr$SNP <- high_pip_eqtls$variant_id
eqtl.gr$gene <- high_pip_eqtls$gene
eqtl.gr$finemap.pip <- high_pip_eqtls$variant_pip
```

Import SNP SNAP results and convert them to hg38 (all eQTL data and ArchR project are in hg38)

```{r}
snp_match <- vroom::vroom('../matched_SNPs/eQTL_top_pip_hg19_5batches/SNPmatch_10_all.txt', col_names = F)
snp_match$X1 <- NULL
snp_match$X2 <- NULL

random.snp.gr.list <- list()
path <- '/project2/gca/software/liftOver_Chains/hg19ToHg38.over.chain'
ch <- import.chain(path)

for(i in 1:ncol(snp_match)){
    curr <- snp_match %>% pull(i)
    chroms <- sapply(strsplit(curr, split = ':'), function(x){x[1]})
    pos <- as.integer(sapply(strsplit(curr, split = ':'), function(x){x[2]}))
    curr.gr <- GRanges(seqnames = chroms, ranges = IRanges(start = pos), SNP = curr)
    seqlevelsStyle(curr.gr) <- "UCSC"
    random.snp.gr.list[[i]] <- unlist(liftOver(curr.gr, ch))
}
```


Get tissue activity of each eQTL from https://zenodo.org/record/3727189 (GTEx v8 paper)

```{r}
eqtl.lfsr <- suppressMessages(vroom::vroom('../eQTL_enrich/broadinstitute-gtex-v8-a014b43/data/Fig6C_all_top.z_lfsr.sig.pruned.txt.gz'))
```

Keep SNPs with PIP > 0.8 AND they have LFSR data.

```{r}
same.snps <- intersect(eqtl.gr$SNP, eqtl.lfsr$variant)
heart.eqtl.lfsr <- eqtl.lfsr[eqtl.lfsr$variant %in% same.snps,]
eqtl.gr <- eqtl.gr[eqtl.gr$SNP %in% same.snps]
saveRDS(eqtl.gr, '../eQTL_enrich/gtex_finemapping/HighPIP_eQTLs_w_LFSR.gr.rds')
```

Find the mean number of tissues per eQTL. 

```{r}
# all eqtls
tissues.active <- rowSums(eqtl.lfsr[,3:ncol(eqtl.lfsr)] < 0.01, na.rm = T) 
all.eqtl.ntissues <- data.frame(eqtl=eqtl.lfsr$variant, ntissues=tissues.active) %>% group_by(eqtl) %>% summarise(mean_tissues = mean(ntissues))
all.eqtl.ntissues <- all.eqtl.ntissues[all.eqtl.ntissues$mean_tissues > 0,]

# heart eqtls
tissues.active <- rowSums(heart.eqtl.lfsr[,3:ncol(heart.eqtl.lfsr)] < 0.01, na.rm = T) 
heart.eqtl.ntissues <- data.frame(eqtl=heart.eqtl.lfsr$variant, ntissues=tissues.active) %>% group_by(eqtl) %>% summarise(mean_tissues = mean(ntissues))
heart.eqtl.ntissues <- heart.eqtl.ntissues[heart.eqtl.ntissues$mean_tissues > 0,]
```

### Tissue Sharing of eQTLs

Below we plot the frequency of tissues active in causal heart eQTLs. Typically eQTLs have a bimodal distribution with peaks at 1-5 tissues and 45-50 tissues. We see that for heart eQTLs, they are generally shared across tissues. 

```{r}
betterHist <- function(X){
 bks <- seq(0, 50, 5)
 labs <- sapply(1:(length(bks)-1), function(x){paste0(bks[x]+1,'-',bks[x+1])})
 bin.count <- table(cut(X$mean_tissues, breaks = bks, labels = labs))
 bin.count <- bin.count/sum(bin.count)
 bin.count.df <- data.frame(count=as.numeric(bin.count), breaks=factor(names(bin.count), levels = labs))
}

all.eqtl.bin <- betterHist(all.eqtl.ntissues)
heart.eqtl.bin <- betterHist(heart.eqtl.ntissues)

result <- rbind(all.eqtl.bin, heart.eqtl.bin)
result$eQTLs <- c(rep(c("All","Heart"), each=nrow(all.eqtl.bin)))
```

```{r}
pdf('../manuscript_figures/figure4/Fig4A_tissue_sharing_eqtl.pdf', width=10, height=8)
ggplot(result, aes(x=breaks, y=count, fill=eQTLs)) + geom_bar(stat="identity", position = "dodge", width=0.8) + ggClean() + xlab('Tissues with LFSR < 0.01') + ylab('Proportion of eQTLs') + ggtitle('Tissue Sharing of Heart eQTLs') + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + scale_fill_manual(values = c("#D6604D","#4393C3"))
dev.off()
```


Add tissue activity to our eQTL genomic ranges object

```{r}
eqtl.gr$ntissues.active <- heart.eqtl.ntissues$mean_tissues[match(eqtl.gr$SNP, heart.eqtl.ntissues$eqtl)]
#eqtl.gr.low <- eqtl.gr[which(eqtl.gr$ntissues.active<=10),]
#eqtl.gr.high <- eqtl.gr[which(eqtl.gr$ntissues.active>=30),]
```

### Distribution of heart eQTLs

Overlap generic genomic annotations with top eQTLs

```{r}
annots <- readRDS('../genomic_annotations/hg38_gtf_genomic_annots.gr.rds')
exon.annots <- annots$exons
utr.annots <- annots$UTRs
exon.annots <- removeOverlaps(X = exon.annots, to.remove = utr.annots)
intron.annots <- annots$introns
```


Split ventrical cardiomyocyte peaks into cell-type specific and into shared.

```{r}
satac <- suppressMessages(ArchR::loadArchRProject('../ArchR/ArchR_heart_latest_noAtrium/', showLogo = F))
union.set <- ArchR::getPeakSet(satac)
union.exons <- union.set[union.set$peakType == "Exonic", ]
union.nonExon <- union.set[union.set$peakType != "Exonic",]
union.nonExon$peakID <- GRToString(union.nonExon)
```
 
How many cell types is each peak from union set found in? Calculcate eQTL enrichment

```{r}
all.calls <- unlist(lapply(list.files('../ArchR/ArchR_heart_latest_noAtrium/PeakCalls', pattern = 'reproduciblePeaks.gr.rds', full.names = T),
                    function(x){readRDS(x)}))

shared.df <- data.frame(peakNum = 1:length(union.set))
shareness <- lapply(all.calls, function(x){unique(queryHits(findOverlaps(union.set, x)))})
shareness <- data.frame(table(unlist(shareness)))
colnames(shareness) <- c("peakNum","freq")
shareness$peakNum <- as.integer(shareness$peakNum)
shared.df <- shared.df %>% left_join(., shareness)

shared.1 <- union.set[shared.df$peakNum[which(shared.df$freq==1)],]
shared.2 <- union.set[shared.df$peakNum[which(shared.df$freq==2)],]
shared.3 <- union.set[shared.df$peakNum[which(shared.df$freq==3)],]
shared.4 <- union.set[shared.df$peakNum[which(shared.df$freq>=4)],]
peaks.gr.list <- list("1"=shared.1,"2"=shared.2,"3"=shared.3,"4"=shared.4)

peaks.eqtls.sharing <- join_overlap_list(gr.list = peaks.gr.list, X = eqtl.gr)
peaks.random.sharing <- lapply(random.snp.gr.list, function(x){
    join_overlap_list(gr.list = peaks.gr.list, X = x)
})

peaks.sharing.type <- lapply(peaks.gr.list, function(x){table(x$peakType)/length(x)})
```

```{r}
random.dist.df.list <- list()
for(i in 1:length(peaks.random.sharing)){
    curr <- peaks.random.sharing[[i]]
    snpsIn <- unique(unlist(sapply(curr, function(x){x$SNP})))
    curr$Unassigned <- random.snp.gr.list[[i]][!(random.snp.gr.list[[i]]$SNP %in% snpsIn),]
    
    curr.df <- as.data.frame(sapply(curr, FUN = function(x){length(x)/length(random.snp.gr.list[[i]])}))
    colnames(curr.df) <- c("freq")
    random.dist.df.list[[i]] <- curr.df
}

random.dist.avg <- data.frame(freq = rowMeans(Reduce(cbind, random.dist.df.list)))
random.dist.avg <- random.dist.avg[1:4,,drop=F]
peak.dist.df <- as.data.frame(sapply(peaks.eqtls.sharing, FUN = function(x){length(x)/length(eqtl.gr)}))
colnames(peak.dist.df) <- c("freq")
```

```{r}
pdf('../manuscript_figures/figure4/FigC_peak_sharing_v_eqtl_propt.pdf', width=8, height=6)
enrichment.df <- data.frame(enrich = c(peak.dist.df$freq,random.dist.avg$freq), numshared = rep(c("1","2","3","4+"), 2), type=rep(c("eQTLs","Random"), each = 4))
ggplot(enrichment.df, aes(x=numshared, y = enrich, fill = type)) + geom_bar(stat = 'identity', position="dodge") + ggClean() + ylab('Proportion of SNPs') + xlab('Peak clusters (# of cell-types containing peak)') 
dev.off()

pdf('../manuscript_figures/figure4/FigC_peak_sharing_v_enrichment.pdf', width=8, height=6)
enrichment.df <- data.frame(enrich = (peak.dist.df$freq/random.dist.avg$freq), numshared = c("1","2","3","4+"))
ggplot(enrichment.df, aes(x=numshared, y = enrich)) + geom_bar(stat = 'identity') + ggClean() + ylab('eQTL Enrichment') + xlab('Peak clusters (# of cell-types containing peak)')
dev.off()
```

eQTL enrichment in cell-type PREs and generic annotations

```{r}
cm.calls <- readRDS('../ArchR/ArchR_heart_latest_noAtrium/PeakCalls/Cardiomyocyte-reproduciblePeaks.gr.rds')
cm.calls <- removeOverlaps(X = cm.calls, to.remove = union.exons)

markers <- readRDS('../ArchR/ArchR_heart_latest_noAtrium/PeakCalls/DA_MARKERS_FDRP_1_log2FC_1.rds')
markers$`Smooth Muscle` <- NULL
markers$Neuronal <- NULL
markers <- lapply(markers, function(x){removeOverlaps(X = x, to.remove = union.exons)})

ct.peaks <- GRToString(unlist(GRangesList(markers)))
shared.peaks <- union.nonExon[!(union.nonExon$peakID %in% ct.peaks),]
non.cm.specific <- union.nonExon[!(union.nonExon$peakID %in% GRToString(markers$Cardiomyocyte)),]
cm.shared <- non.cm.specific[unique(queryHits(findOverlaps(non.cm.specific, cm.calls))),]

intron.nonpeak <- removeOverlaps(intron.annots, to.remove = union.set)
```


Overlap eQTLs with exons, introns, etc.

```{r}

peaks.gr.list <- c(markers, 
                   list("CM Shared" = cm.shared,
                        "Shared" = shared.peaks,
                        "Intronic" = intron.nonpeak,
                        "Exon" = exon.annots, 
                        "UTR" = utr.annots))

peaks.eqtls <- join_overlap_list(gr.list = peaks.gr.list, X = eqtl.gr)
peaks.random <- lapply(random.snp.gr.list, function(x){
    join_overlap_list(gr.list = peaks.gr.list, X = x)
})
```

```{r}
random.dist.df.list <- list()
for(i in 1:length(peaks.random)){
    curr <- peaks.random[[i]]
    snpsIn <- unique(unlist(sapply(curr, function(x){x$SNP})))
    curr$Unassigned <- random.snp.gr.list[[i]][!(random.snp.gr.list[[i]]$SNP %in% snpsIn),]
    
    curr.df <- as.data.frame(sapply(curr, FUN = function(x){length(x)/length(random.snp.gr.list[[i]])}))
    colnames(curr.df) <- c("freq")
    random.dist.df.list[[i]] <- curr.df
}
```

```{r}
random.dist.avg <- data.frame(freq = rowMeans(Reduce(cbind, random.dist.df.list)))
random.dist.avg$category <- rownames(random.dist.avg)
```

```{r}
snpsIn <- unique(
  unlist(
    sapply(peaks.eqtls, function(x){x$SNP})
  )
)
peaks.eqtls$Unassigned <- eqtl.gr[!(eqtl.gr$SNP %in% snpsIn),]
```


```{r}
peak.dist.df <- as.data.frame(sapply(peaks.eqtls, FUN = function(x){length(x)/length(eqtl.gr)}))
colnames(peak.dist.df) <- c("freq")
peak.dist.df$category <- rownames(peak.dist.df)
```

```{r}
peak.set.dist.df <- Reduce(rbind, list(random.dist.avg, peak.dist.df))
peak.set.dist.df$SNPs <- c(rep("Random SNPs", nrow(random.dist.avg)),
                          rep("eQTLs", nrow(peak.dist.df)))
```


```{r}
plot.levels <- c("Exon","UTR","Intronic",names(markers),"CM Shared","Shared", "Unassigned")
#plot.levels <- c(names(union.set.split), "Unassigned")
peak.set.dist.df$category <- factor(peak.set.dist.df$category, 
                                    levels = plot.levels)
#peak.set.dist.df <- peak.set.dist.df[peak.set.dist.df$category!="Unassigned",]

n_cat <- nrow(peak.dist.df)
# enrich <- p.adjust(sapply(1:n_cat, function(y){binom.test(x = round(1216*peak.dist.df$freq[y]), n = 1216, p = random.dist.avg$freq[y], alternative = "greater")$p.value}), method = 'bonferroni')
enrich <- peak.dist.df$freq / random.dist.avg$freq
enrich.df <- data.frame(category= peak.dist.df$category,
                        group1 = rep("eQTLs", n_cat),
                        group2 = rep("Random SNPs", n_cat),
                        p = round(enrich,2),
                        y.position = sapply(1:n_cat, function(x)max(peak.dist.df$freq[x], random.dist.avg$freq[x]))) %>% 
    as_tibble() %>% 
    mutate(group1 = unfactor(group1), group2 = unfactor(group2))
enrich.df$category <- factor(enrich.df$category, levels = plot.levels)
enrich.df <- enrich.df %>% add_x_position(x = "category", dodge = 0.75)
enrich.df$p[enrich.df$p == 0] <- "ns"

#pdf('../manuscript_figures/figure4/Fig4B_wc_union_set_distribution.pdf', width=6, height=4)
ggbarplot(peak.set.dist.df, x = "category", y = "freq", fill = "SNPs", position = position_dodge(.75), palette = "Paired") + 
    stat_pvalue_manual(enrich.df, tip.length = 0.01, step.increase = 0, bracket.nudge.y = 0.01) + coord_cartesian(ylim = c(0, 0.45)) +
    ggClean(rotate_axis = T)

#dev.off()
```


Tissue sharing patterns in scATAC-seq peak sets

```{r}
peaks.eqtls$Unassigned <- NULL
peaks.tissue.sharing <- sapply(peaks.eqtls, function(x){x$ntissues.active}) 
peaks.tissue.sharing$`Union Set` <- NULL
peaks.tissue.sharing <- setNames(unlist(peaks.tissue.sharing, use.names=F),rep(names(peaks.tissue.sharing), lengths(peaks.tissue.sharing)))
peaks.tissue.sharing.df <- data.frame(category=names(peaks.tissue.sharing), ntissues=peaks.tissue.sharing)
peaks.tissue.sharing.df$category <- factor(peaks.tissue.sharing.df$category, levels=plot.levels)
peaks.tissue.sharing.df <- peaks.tissue.sharing.df[!is.na(peaks.tissue.sharing.df$ntissues),]

peaks.tissue.sharing.df$category <- factor(peaks.tissue.sharing.df$category, levels = plot.levels)
```

```{r}
pdf("../manuscript_figures/figure4/Fig4B_tissues_active_dist.pdf", width=15, height=6)

ggplot(peaks.tissue.sharing.df, aes(x=category, y=ntissues)) + 
  geom_violin(fill="lightblue") + 
  geom_boxplot(width=0.1) + 
  ggClean() + LegendOff() + xlab("Tissues Active (LFSR < 0.01)") + ylab("") +
  scale_y_continuous(breaks=seq(0,50,10))

dev.off()
```

```{r}
diff.exp.res <- readRDS('../seurat/diff_expr_markers.df.rds')
cluster.gene <- diff.exp.res[,c("cluster", "gene")]
```


### eGene Expression

```{r}
srna <- readRDS('../seurat/Heart_RNA_Processed_Combined_NoAtrium.rds')
srna.exp <- Seurat::AverageExpression(srna)
srna.exp.mat <- as.matrix(log2(srna.exp$RNA + 1))
agg.exp <- as.matrix(log2(rowMeans(srna.exp$RNA) + 1))
colnames(agg.exp) <- "Bulk"
```

```{r}
eqtl.snp.gr <- readRDS('../eQTL_enrich/gtex_finemapping/HighPIP_eQTLs_w_LFSR.gr.rds')

finemap.res <- readr::read_tsv( '../eQTL_enrich/gtex_finemapping/GTEx_v8_finemapping_DAPG/Heart_LV_Finemapping_CS95.txt', col_names = F)
colnames(finemap.res) <- c("tissue","gene","cluster_id","cluster_pip","variant_id","variant_pip")
finemap.res <- finemap.res[finemap.res$variant_id %in% eqtl.snp.gr$SNP,] %>% 
    dplyr::select(variant_id, gene) %>% rename(gene_id = gene, SNP = variant_id) %>% mutate(gene_id = sub('\\..*', '', gene_id))

ensembl.to.symbol <- AnnotationDbi::select(x = org.Hs.eg.db::org.Hs.eg.db, keys = finemap.res$gene_id, columns = "SYMBOL", keytype = "ENSEMBL") %>% 
    rename(gene_id = ENSEMBL) %>%
    group_by(gene_id) %>% slice(1)

finemap.res <- left_join(finemap.res, ensembl.to.symbol)

finemap.res <- finemap.res[finemap.res$SYMBOL %in% rownames(srna.exp.mat),]
```

```{r}
finemap.res.rna <- finemap.res %>% rename(gene = SYMBOL) %>% left_join(., cluster.gene)
finemap.res.rna <- finemap.res.rna[!(finemap.res.rna$cluster %in% c("Smooth Muscle","Neuronal")),]
tissues.active.df <- data.frame(SNP = finemap.res.rna$SNP, 
                                tissues.active = eqtl.gr$ntissues.active[match(finemap.res.rna$SNP, eqtl.gr$SNP)], 
                                cluster = unfactor(finemap.res.rna$cluster),
                                stringsAsFactors = F)
tissues.active.df$cluster[is.na(tissues.active.df$cluster)] <- "Shared"
tissues.active.df$cluster <- factor(tissues.active.df$cluster, levels = rev(c("Cardiomyocyte","Pericyte","Endothelial","Fibroblast","Lymphoid","Myeloid","Shared")))

pdf('../manuscript_figures/figure4/eGene_RNA_sharing_violin.pdf', width=5, height=8)
ggplot(tissues.active.df, aes(x = cluster, y = tissues.active)) + geom_violin(fill='lightblue') + geom_boxplot(width = 0.1) + ggClean(rotate_axis = T) + xlab('Tissue Activity of eQTL') + ylab('eGene Type') + coord_flip()
dev.off()
```

```{r}
prop.exp <- 0.5
rna.binary.exp <- (srna@assays$RNA@counts>0)
same.genes <- intersect(rownames(srna.exp.mat), finemap.res$SYMBOL)

celltypes <- unique(Seurat::Idents(srna))
celltype.prop.exp <- list()
for(ct in celltypes){
    celltype.prop.exp[[ct]] <- rowMeans(rna.binary.exp[,Seurat::Idents(srna) == ct])
}
rna.prop.exp.mat <- Reduce(cbind, celltype.prop.exp)
colnames(rna.prop.exp.mat) <- celltypes

egene.ncelltypes <- rowSums(rna.prop.exp.mat[same.genes,] >= prop.exp)
egene.ncelltypes <- egene.ncelltypes[egene.ncelltypes>0]
egene.ncelltypes[egene.ncelltypes >=4] <- 4
prop.egenes.in <- table(egene.ncelltypes)/length(same.genes)

random.ncelltypes <- rowSums(rna.prop.exp.mat[sample(1:nrow(rna.prop.exp.mat), size = length(same.genes), replace = F),] >= prop.exp)
random.ncelltypes <- random.ncelltypes[random.ncelltypes>0]
random.ncelltypes[random.ncelltypes>=4] <- 4
prop.random.in <- table(random.ncelltypes)/length(same.genes)

barplot(prop.egenes.in/prop.random.in)
```



```{r}
celltype_ideal_order <- c("Cardiomyocyte","Smooth Muscle","Pericyte","Endothelial","Fibroblast","Neuronal", "Lymphoid","Myeloid")

ns <- rep(names(peaks.eqtls), lengths(peaks.eqtls))
eqtl.overlap.peaks <- unlist(GRangesList(peaks.eqtls))
eqtl.overlap.peaks$cellType <- factor(ns, levels = c("Cardiomyocyte","CM Shared")) # only picking these two
eqtl.overlap.peaks.df <- eqtl.overlap.peaks %>% as_tibble() %>% select(SNP, cellType)

finemap.res.2 <- finemap.res %>% left_join(., eqtl.overlap.peaks.df, on='SNP')

finemap.res.inRNA <- finemap.res.2[finemap.res.2$SYMBOL %in% rownames(srna.exp.mat),] %>% arrange(cellType) %>% filter(!is.na(cellType))
exp.df <- data.frame(log2cpm = c(srna.exp.mat[finemap.res.inRNA$SYMBOL,"Cardiomyocyte"]), type = finemap.res.inRNA$cellType)

ggplot(exp.df, aes(x = type, y = log2cpm)) + geom_boxplot(width=0.2) + ggClean() + xlab('eGene Type') + ylab('log2 TP10k') + 
    annotate(geom="text", x=2, y=4, label="t-test p-value: 0.0002",color="red")
```

```{r}
ns <- rep(names(peaks.eqtls.sharing), lengths(peaks.eqtls.sharing))
eqtl.overlap.peaks <- unlist(GRangesList(peaks.eqtls.sharing))
eqtl.overlap.peaks$cellType <- factor(ns, levels = 1:4) # only picking these two
eqtl.overlap.peaks.df <- eqtl.overlap.peaks %>% as_tibble() %>% select(SNP, cellType)

finemap.res.2 <- finemap.res %>% left_join(., eqtl.overlap.peaks.df, on='SNP')

finemap.res.inRNA <- finemap.res.2[finemap.res.2$SYMBOL %in% rownames(srna.exp.mat),] %>% arrange(cellType) %>% filter(!is.na(cellType))
exp.df <- data.frame(log2cpm = c(agg.exp[finemap.res.inRNA$SYMBOL,"Bulk"]), type = finemap.res.inRNA$cellType)

ggplot(exp.df, aes(x = type, y = log2cpm)) + geom_boxplot(width=0.2) + ggClean() + xlab('Number of Cell-types Shared') + ylab('eGene Bulk Expression') + ylim(c(0,1))
```


### overlap with iQTLs

```{r}
iqtls <- readr::read_tsv('../iQTLs_summary_stats.txt')
iqtls <- iqtls[iqtls$celltype=="Myocytes" & iqtls$tissue_id=="Heart_Left_Ventricle",]
iqtls.gr <- snpIDtoGR(iqtls$variant_id)
```

```{r}
markers <- readRDS('../ArchR/ArchR_heart_latest_noAtrium/PeakCalls/DA_MARKERS_FDRP_1_log2FC_1.rds')
markers$`Smooth Muscle` <- NULL
markers$Neuronal <- NULL

peaks.iqtls <- join_overlap_list(gr.list = markers, X = iqtls.gr)
peak.iqtls.dist.df <- as.data.frame(sapply(peaks.iqtls, FUN = function(x){length(x)/length(iqtls.gr)}))
colnames(peak.iqtls.dist.df) <- c("freq")
peak.iqtls.dist.df$category <- rownames(peak.iqtls.dist.df)
```


### Sharing in non CMs

The main issue to address is why eQTLs in non-CM peaks are shared with other tissues at high proportions. The hypothesis is that some of the cell types, say fibroblast, are commonly present in other tissues, which drive the high sharing.
So if the hypothesis is true, we expect that tissues with shared cell types would have high sharing (in non-CM eQTLs), while tissues not shared cell types would have low sharing.


```{r}
dapg.eqtl <- readr::read_tsv(file = '../eQTL_enrich/gtex_finemapping/GTEx_v8_finemapping_DAPG/GTEx_v8_finemapping_DAPG.txt.gz')
```

```{r}
peaks.eqtls.atac <- list(peaks.eqtls$Cardiomyocyte, 
                         peaks.eqtls$Fibroblast, 
                         peaks.eqtls$Endothelial,
                         peaks.eqtls$Lymphoid)
names(peaks.eqtls.atac) <- c("Cardiomyocyte","Fibroblast","Endothelial","Lymphoid")
tissues.to.keep <- c("Whole_Blood","Testis","Brain_Amygdala", "Cells_Cultured_fibroblasts","Heart_Atrial_Appendage","Muscle_Skeletal")
dapg.eqtl.filt <- dapg.eqtl %>% 
    filter(variant_pip > 0.8) %>%
    filter(tissue_id %in% tissues.to.keep) %>% 
    mutate(eqtl_identifier = paste0(variant_id, '__', gene_id))

res <- lapply(peaks.eqtls.atac, function(x){
    eqtlsIn <- x %>% as_tibble() %>% mutate(eqtl_identifier = paste0(SNP, '__', gene)) %>% .$eqtl_identifier;
    dapg.eqtl.filt %>% group_by(tissue_id) %>% summarise(non_cm_overlap = sum(eqtlsIn %in% eqtl_identifier)/length(eqtlsIn))
})
res.mat <- Reduce(rbind, res)
res.mat$type <- rep(names(res), each=length(tissues.to.keep))
``` 

```{r}
pdf('../manuscript_figures/figure4/Fig4G_eQTLs_similar_tissues.pdf', width = 6, height=5)
ggplot(res.mat, aes(x= tissue_id, y = non_cm_overlap, fill=type)) + geom_bar(stat='identity', position='dodge') + ggClean(rotate_axis = F) +
    xlab('Proportion of eQTLs') + ylab('Tissue') +   
    scale_fill_manual(values = c("#b22222","#8DD3C7","#BEBADA", "#FB8072")) + LegendOff() + scale_y_continuous(breaks = seq(0, 0.55, 0.1)) +
    coord_flip()
dev.off()
```

