---
title: "QC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Seurat)
library(ArchR)
library(ggplot2)
library(dplyr)
source('../R/analysis_utils.R')
```

### scRNA-seq

```{r}
srna <- readRDS('../seurat/Heart_Seurat_RNA_all_samples_raw.RDS')
```

```{r}
donors <- c("02207","02336","03231")
qc.df <- data.frame(umi = srna$nCount_RNA, Donor=srna$individual)
df.list <- list()
for(d in donors){
  curr.df <- qc.df %>% filter(Donor == d) %>% arrange(-umi) %>% mutate(rank = 1:nrow(.)) %>% mutate(Type=ifelse(umi > 1000, 'Cell','Background'))
  df.list[[d]] <- curr.df
}
qc.df <- Reduce(rbind, df.list)

```

#### Barcode-rank plot

This plot shows each barcode's UMI count and rank. It is useful for distinguishing actual cells vs background.

```{r}
ggplot(qc.df, aes(x = rank, y=umi, color=Donor)) + geom_line(aes(linetype = Type)) + ggClean() +
  scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_linetype_manual(values = c("dashed","solid")) + 
  xlab('Barcodes') + ylab('UMI')

```

#### Features vs UMI count

This plot shows how many genes are detected at each level of UMI count. We use UMI count as a proxy for read count which we do not record directly. This plot is useful for setting lower and upper limits on features as to filter doublets and background barcodes. 

```{r}
qc.df <- data.frame(umi = srna$nCount_RNA/10^4, features=srna$nFeature_RNA/1000, Donor=srna$individual)
ggplot(qc.df, aes(x=umi, y=features, color=Donor)) + geom_point(alpha=0.5) + ggClean() + xlab(expression('UMI ('%*% 10^4*')')) + ylab(expression('Features (' %*% 10^3*')'))
```

### scATAC-seq

```{r}
d22 <- loadArchRProject('../ArchR_project_02207/', showLogo = F)
d23 <- loadArchRProject('../ArchR_project_02336/', showLogo = F)
d32 <- loadArchRProject('../ArchR_project_03231/', showLogo = F)
```

#### Library size distribution

```{r}
comb <- suppressMessages(loadArchRProject('../ArchR/ArchR_heart/', showLogo = F))
```

```{r}
ATAC_SAMPLES <- c("MW200804AA","MW200804AB","MW200804AC","MW200804AD","SP-HE-MW200928E1ATAC-411ATAC","SP-HE-MW200928E1ATAC-413ATAC","SP-HE-HE200915ATAC-175ATAC","SP-HE-HE200915ATAC-359ATAC","SP-HE-HE200915ATAC-360ATAC","SP-HE-HE200915ATAC-397ATAC","SP-HE-MW200928E1ATAC-396ATAC","SP-HE-MW200928E1ATAC-398ATAC","SP-HE-MW200928E2ATAC-175ATAC","SP-HE-MW200928E2ATAC-366ATAC","SP-HE-MW200928E2ATAC-367ATAC","SP-HE-MW200928E2ATAC-407ATAC","SP-HE-MW200928E1ATAC-406ATAC","SP-HE-MW200928E1ATAC-408ATAC")
ATAC_INDIVIDUALS <- c(rep("black",6),rep("blue",6),rep("red",6))

curr_x <- sort(log2(comb$nFrags[comb$Sample==ATAC_SAMPLES[1]]))
curr_y <- 1:length(curr_x)/length(curr_x)
plot(curr_x, curr_y, type='l', lwd=3, col=ATAC_INDIVIDUALS[15], xlab='log2 nFrags', ylab='CDF', xlim=c(11, 17))

for(i in 2:length(ATAC_SAMPLES)){
  
  curr_x <- sort(log2(comb$nFrags[comb$Sample==ATAC_SAMPLES[i]]))
  curr_y <- 1:length(curr_x)/length(curr_x)
  lines(curr_x, curr_y, lwd=3, col=ATAC_INDIVIDUALS[i])
  
}

legend(x = 16, y = 0.3, legend = c("02207","02336","03231"), pch = 16, col=c("black","blue","red"), bty = "n")
```


#### TSS Enrichment

This shows the distribution of TSS enrichment across cells. Cells with low value of TSS enrichment are low quality. 

```{r}
tss.df <- data.frame(tss.enrich = c(d22$TSSEnrichment, d23$TSSEnrichment, d32$TSSEnrichment), 
                     Donor = c(rep('02207', length(d22$cellNames)), rep('02336', length(d23$cellNames)), rep('03231', length(d32$cellNames))))
```

```{r}
ggplot(tss.df) + stat_density(aes(x=tss.enrich, color=Donor), geom="line", position="identity", size=1.5) + ggClean() + xlab('TSS Enrichment') + 
    theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

#### Fragment size distribution

This plot tells us the fragment size distribution. A good ATAC-seq library will have two peaks: one around ~50 bp representing the pitch of DNA and one around ~150 representing actual open chromatin peaks. 

```{r}
n_down_sample <- 10^6

d22.sizes <- readRDS('../ArchR_project_02207/fragment_sizes.rds')
d22.sizes <- d22.sizes[sample(1:length(d22.sizes), n_down_sample, replace=F)]

d23.sizes <- readRDS('../ArchR_project_02336/fragment_sizes.rds')
d23.sizes <- d23.sizes[sample(1:length(d23.sizes), n_down_sample, replace=F)]

d32.sizes <- readRDS('../ArchR_project_03231/fragment_sizes.rds')
d32.sizes <- d32.sizes[sample(1:length(d32.sizes), n_down_sample, replace=F)]

```

```{r}
size.df <- data.frame(sizes=c(d22.sizes, d23.sizes, d32.sizes), Donor=c(rep("02207",length(d22.sizes)), rep("02336",length(d23.sizes)), rep("03231", length(d32.sizes))))
size.df <- size.df[size.df$sizes < 1000, ]
```

```{r}
ggplot(size.df) + stat_density(aes(x=sizes, color=Donor), geom="line", position="identity", size=1.5) + ggClean() + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  xlab('Fragment Size')
```

